Between Brussels and Britain: Uncovering the driving forces of Brexit

### 2024-04-19 paper/paper.qmd
Finishing up some sections

---
title: "Between Brussels and Britain: Uncovering the driving forces of Brexit"
subtitle: "A statistical approach to assessing engagement in a supranational union"
author: 
  - Samarth Rajani
thanks: "Code and data are available at: https://github.com/samarthrajani/-Between-Brussels-and-Britain-Understanding-Party-Engagement-with-EU-and-Domestic-Topics-in-the-UK-."
date: today
date-format: long
abstract: "In a union, all parties must work together to guarantee best outcomes. In the case of the United Kingdom, friction between legislative bodies within the nation and the European Parliament caused the UK to exit the union. This paper looks at a speech based dataset from the UK House of Commons to suggest which political parties were more vocal in the build up to Brexit, giving insight into which political parties were the driving force in this change.  "
format: pdf
number-sections: true
bibliography: references.bib
---

@@

Keeping this in mind, it is worth examining how often matter of the European Union dictated discourse in the United Kingdom. Specifically, this paper studies how often each party gave speeches that involved discussions around the European Union right down to the level of individual sentences. Looking at the ParlEE(@dataset) dataset, this paper examines how likely is one political party to devote time to discussing EU related matters as compared to domestic ones. S

\[Paragraph on results\]

The remainder of this paper is structured as follows. @sec-data....

@@

# Model

@@

# Results

@@

summary(first_model)
```

# Discussion

## First discussion point {#sec-first-point}

@@

When studying the relationship between two variables, there is a concern of reverse causality. For example, if we were to study the impact on institutions on the GDP of a nation,

\appendix

# Appendix {.unnumbered}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

#plot(first_model, "trace")

#plot(first_model, "rhat")
```

>>>>

---
title: "Between Brussels and Britain: Uncovering the driving forces of Brexit"
subtitle: "A statistical approach to assessing engagement in a supranational union"
author: 
  - Samarth Rajani
thanks: "Code and data are available at: https://github.com/samarthrajani/-Between-Brussels-and-Britain-Uncovering-the-driving-forces-of-Brexit/tree/main"
date: today
date-format: long
abstract: "In a union, all parties must work together to guarantee best outcomes. In the case of the United Kingdom, friction between legislative bodies within the nation and the European Parliament caused the UK to exit the union. This paper looks at a speech based dataset from the UK House of Commons to suggest which political parties were more vocal in the build up to Brexit, giving insight into which political parties were the driving force in this change.  "
format: pdf
number-sections: true
bibliography: references.bib
---

@@

Keeping this in mind, it is worth examining how often matter of the European Union dictated discourse in the United Kingdom. Specifically, this paper studies how often each party gave speeches that involved discussions around the European Union right down to the level of individual sentences. Looking at the ParlEE(@dataset) dataset, this paper examines how likely is one political party to devote time to discussing EU related matters. It is a reasonable assumption to make that party ideaologies that significantly deviate from and oppose the European Union will cause those parties to be more vocal about EU related matters, thereby increasing the incidence of EU related discourse. This paper utilises this assumption to study major contributors to the Brexit in 2020.

The results reflect support for Brexit reasonably well. The UKIP and GPEW very discussing EU related matters most often, while the Liberal Democrats who oppose Brexit discussed EU matters the least.

The remainder of this paper is structured as follows. @sec-data which discusses the dataset in high detail, @sec-model which outlines the model used for analysis and lastly, @sec-results which explains the results from the model. There is a discussion section @sec-discussion which developes ideas for future studies, the strengths and the weaknesses of this paper.

@@

# Model {#sec-model}

@@

# Results {#sec-results}

@@

#summary_result <- summary(first_model)

coefficients <- coef(first_model)
# Extract variable names (excluding intercept)
variables <- names(coefficients)[-1]  # Exclude the intercept

# Extract coefficients (excluding intercept)
coefficients <- coefficients[-1]  # Exclude the intercept

# Create data frame
coefs_df <- data.frame(Party = variables, Coefficients = coefficients)

# Print the data frame
print(coefs_df)
```

The summary table provides the coefficients for our regression.

The reference category for this regression is the The Alliance Party of Northern Ireland (APNI). All the coefficients show how more or less likely is each party to mention the European Union in their speeches in log odds. For example, the conservative party is 1.2 log odds more likely to discuss the EU in their speech as compared to APNI.

Our model seems to be a good predictor for support for Brexit. The UK Independence Party (UKIP) has most strongly favoured Brexit according to @brexitparty, which as can be seen has a 1.7 log odds coefficient which is the second largest. The highest coefficient is held by the Green Party of England and Wales (GPEW), which also held pro-Brexit views until 2019, which is in line with our model timeframe.

Furthermore, the Liberal Democrat party (LibDem) have been against Brexit, beieving that being a part of the European Union is in the best interests of the UK government. This can be seen as a coefficient of 0.8 log odds is one of the lowest in the summary table.

It is worth stating that the reference ctaegory variable here is Alliance Party of Northern Ireland (APNI), which has holds an anti-Brexit stance.

# Discussion {#sec-discussion}

## First discussion point

@@

When studying the relationship between two variables, there is a concern of reverse causality. For example, if we were to study the impact on institutions on the GDP of a nation, our estimates would be biased since GDP could also affect and improve an institituion. To tackle this, an Instrumental Variable approach can be used.

For this to happen,
